---
title: "001_netcdf_data_"
author: "Deidre Jaeger"
date: "1/7/2019"
output: html_document
---

```{r install-packages}
library(ncdf4)
library(RColorBrewer)
library(lattice)
library(raster)
library(rasterVis)
library(proj4)
#install.packages("proj4")
library(gridExtra)
library(rgeos)
library(rgdal) # needed to read in boulder shapefile
library(daymetr)
```

```{r open-netcdf-tmin}
dir <- "/Users/deidrejaeger/Documents/Career/CU-Boulder/Research/BoulderMicroclimate/boulder-microclimate-analysis/data/daymet/V3/CF_tarred/tars_2016/"
year <- "11738_2016/"
ncname <- "tmin"
ncfname <- paste(dir, year, ncname, ".nc", sep="")
ncin <- nc_open(ncfname, write=F)
print(ncin)
```
# there are missing values that are coded as -9999, we need to tell R to read those as NA

```{r get x-y-variables}
x <- ncvar_get(ncin, "x")
y <- ncvar_get(ncin,"y")
yearday <- ncvar_get(ncin,"yearday")
nc_close(ncin)

```

```{r function-to-retrieve-variables}
getNCDFClimate <- function(var, year){ # Two arguments define the climate variable and year of the netCDF
ncfname <- paste(dir, year, ncname, ".nc", sep="")
ncin <- nc_open(ncfname, write=T)
var.array <- ncvar_get(ncin,var) # get the climate variable values
nc_close(ncin)
var.vec.long <- as.vector(var.array) # convert from array to vector
return(var.vec.long)
}

```

```{r get-tmin-vals}
vec.tmin <- getNCDFClimate("tmin", "11738_2016/")
```

```{r create-matrix}
ncells <- dim(x)*dim(y)
nyearday <- dim(yearday)
tmin.mat <- matrix(vec.tmin, nrow=ncells, ncol=nyearday, byrow=F)

```

```{r plotting}
grid <- expand.grid(xlon=x, ylat=y)
xyz.tmin <- cbind(grid, tmin.mat[,92])
head(xyz.tmin)
```
Looking at the xyz.tmin values, it seems they are ranging between -10 to -1.5 degrees C, or 3 to 34 degrees F

```{r construct-proj-and-raster}
pj <- "+proj=lcc +lat_1=25.00 +lat_2=60.00 +lat_0=42.5 +lon_0=-100.00 +x_0=0 +y_0=0 +ellps=GRS80 +units=m +no_defs"
rast.tmin <- rasterFromXYZ(xyz.tmin, crs=pj)

# plot
cuttmp <- c(-30,-20,-10,-5,-2,0,2,5,10,20,30) # define break points for color scale
colkey <- list(at=cuttmp, labels=list(at=cuttmp)) # colorkey helps put legend labels at correct break points
p1 <- levelplot(rast.tmin, margin=FALSE, colorkey=colkey, at=cuttmp, cuts=12, col.regions=rev(brewer.pal(11,"RdBu")), main=list(label="Min Temperature (April 1, 1980)", cex=.8), scales=list(draw=FALSE))

grid.arrange(p1, ncol=1)
```

```{r load-boulder-city-outline}
boulder_outline <- readOGR("data/BoulderCityLimits/BoulderCityLimits.shp")
plot(boulder_outline)
extent(boulder_outline)
crs(boulder_outline)
```

```{r reproject-raster-to-feet}
rast.tmin_reproj <- projectRaster(rast.tmin, crs="+proj=lcc +lat_1=25.00 +lat_2=60.00 +lat_0=42.5 +lon_0=-100.00 +x_0=0 +y_0=0 +ellps=GRS80 +units=us-ft +no_defs")
crs(rast.tmin_reproj)


# Reproject the city polygons onto the raster brick's coord ref system
boulder_outline_reproj <- spTransform(boulder_outline,
                                crs(rast.tmin_reproj))
```


```{r clip-raster-to-boulder}
# trim extent of raster
tmin.rast.crop <- crop(rast.tmin_reproj, boulder_outline_reproj)
plot(tmin.rast.crop)
plot(boulder_outline_reproj, add = TRUE)

```
Ok- finally back to the same point I was at before the debackle of downloading ggspatial


Still have the odd values problem

```{r 2016-data}
# create a region of interest, example is: c(top left lat, top left long, bottom right lat, bottom right long)
roi <- c(40.09, -105.32, 39.95, -105.17)
# Define the variables for data extraction
start_year <- 2016
end_year <- 2016
climate_var <- "tmin" # c("tmin", "tmax") or "ALL"
data_path <- "data/daymet/boulder-roi-40.11-105.34"
data_path2 <- "/Volumes/UrbanClimatePheno/daymet" # i think the file is too large for my computer, switch to an external harddrive for raw data

# use daymet API to download the data in the region of interest
download_daymet_tiles(location = roi,
                      tiles = NULL,
                      start = start_year,
                      end = end_year,
                      param = climate_var,
                      path = data_path2, 
                      silent = F)

tile1 <- "/tmin_2016_11558"
tmin.s <- raster::stack(paste0(data_path2,tile1,".nc"))

plot(tmin.s)

tmin_s <- raster::stack(tmin_r)
raster::plot(tmin_s)
tmin.b <- raster::brick(tmin_s)
extent(tmin.b)
crs(tmin.b)
plot(tmin.b)

tmin.b@data


tmin_r$daily.minimum.temperature
```

```{r 2012-monthly-data}
# create a region of interest, example is: c(top left lat, top left long, bottom right lat, bottom right long)
roi <- c(40.097, -105.304, 40.002, -105.175)
# Define the variables for data extraction
start_year <- 2012
end_year <- 2012
climate_var <- "tmin" # c("tmin", "tmax") or "ALL"
data_path <- "data/daymet/boulder-roi-40.097-105.304"
data_path2 <- "/Volumes/UrbanClimatePheno/daymet" # i think the file is too large for my computer, switch to an external harddrive for raw data

# use daymet API to download the data in the region of interest
download_daymet_ncss(location = roi,
                      start = start_year,
                      end = end_year,
                      frequency = "monthly",
                      param = climate_var,
                      path = data_path2, 
                      silent = F)
tmin.mon.s <- raster::stack(paste0(data_path2,"/tmin_monavg_2012_ncss.nc"))

print(tmin.mon.s)
plot(tmin.mon.s)
plot(tmin.mon.s[[2]])

crs(tmin.mon.s)
extent(tmin.mon.s)

# Reproject the city polygons onto the raster brick's coord ref system
boulder_outline_reproj2 <- spTransform(boulder_outline,
                                crs(tmin.mon.s))
crs(boulder_outline_reproj2)
extent(boulder_outline_reproj2)
plot(boulder_outline_reproj2)
plot(tmin.mon.s[[2]], add = T)

# try reprojecting
tmin.mon.s_repr <- projectRaster(tmin.mon.s, crs="+proj=lcc +lat_1=25.00 +lat_2=60.00 +lat_0=42.5 +lon_0=-100.00 +x_0=0 +y_0=0 +ellps=GRS80 +units=m +no_defs")

plot(tmin.mon.s_repr)

# Reproject the city polygons onto the raster brick's coord ref system
boulder_outline_reproj3 <- spTransform(boulder_outline,
                                crs(proj4string(tmin.mon.s_repr)))
extent(tmin.mon.s_repr)
extent(boulder_outline_reproj3)

plot(boulder_outline_reproj3)
plot(tmin.mon.s_repr[[2]], add = T)

plot(tmin.mon.s_repr[[2]])
plot(boulder_outline_reproj3, add = T)

```


```{r 2013-daily}
# create a region of interest, example is: c(top left lat, top left long, bottom right lat, bottom right long)
roi <- c(40.09, -105.32, 39.95, -105.17)
# Define the variables for data extraction
start_year <- 2013
end_year <- 2013
climate_var <- "tmin" # c("tmin", "tmax") or "ALL"
data_path2 <- "/Volumes/UrbanClimatePheno/daymet" # i think the file is too large for my computer, switch to an external harddrive for raw data

# use daymet API to download the data in the region of interest
download_daymet_tiles(location = roi,
                      start = start_year,
                      end = end_year,
                      tile = NULL,
                      param = climate_var,
                      path = data_path2, 
                      silent = F)
tmin2.s <- raster::stack(paste0(data_path2,"/tmin_2012_11738.nc"))
plot(tmin2.s)

crs(tmin2.s)

# Reproject the city polygons onto the raster brick's coord ref system
boulder_outline_reproj4 <- spTransform(boulder_outline,
                                crs(proj4string(tmin2.s)))
crs(boulder_outline_reproj4)
extent(boulder_outline_reproj4)
extent(tmin2.s)


# try reprojecting
tmin.mon.s_repr <- projectRaster(tmin.mon.s, crs="+proj=lcc +lat_1=25.00 +lat_2=60.00 +lat_0=42.5 +lon_0=-100.00 +x_0=0 +y_0=0 +ellps=GRS80 +units=m +no_defs")

plot(tmin.mon.s_repr)

# Reproject the city polygons onto the raster brick's coord ref system
boulder_outline_reproj3 <- spTransform(boulder_outline,
                                crs(proj4string(tmin.mon.s_repr)))


plot(tmin2.s[[2]])
plot(boulder_outline_reproj3, add = TRUE)



# trim extent of raster
tmin2.s.crop <- crop(tmin2.s, boulder_outline_reproj3)

plot(tmin2.s.crop[[2]])
plot(boulder_outline_reproj3, add = T)

plot(tmin2.s.crop[[100]])
plot(boulder_outline_reproj3, add = T)

crs(boulder_outline_reproj3)
extent(boulder_outline_reproj3)

```

crs(boulder_outline_reproj3)

CRS arguments:
 +proj=lcc +lat_1=25.00 +lat_2=60.00 +lat_0=42.5 +lon_0=-100.00 +x_0=0 +y_0=0 +ellps=GRS80 +units=m
+no_defs 

extent(boulder_outline_reproj3)
> extent(boulder_outline_reproj3)
class       : Extent 
xmin        : -431996.4 
xmax        : -421564.2 
ymin        : -255924.2 
ymax        : -241673.5 


