---
title: "02_bmc_data_exploration_daily"
author: "Deidre Jaeger"
date: "1/9/2019"
output: html_document
---

Exploration plan : Start with tmin 2013

Define a study radius around the city
convert to F

1) maps of first 9 days of Jan, Apr, Jul, Oct (DONE)

2) mean min temp maps for each month for a cell 

3) calculate the coefficient of variation (sd/mean * 100) for each month and year for a cell

4) Look at yearly averages of tmin for a cell from 1980 to 2017 (tmin and tmax)

Questions:
Is there a spatial pattern to temp variability? 
Is there a detectable urban heat island effect? 
How does this compare to the overall climate change signal? (using landsat?)

Analysis plan
1) define an area in the city and outside, compare the mean yearly temps for a section inside the city to a section outside the city see if significantly different for a 30 year comparison- what is the difference?
?how many weather stations are in boulder and where are they? 

```{r install-packages}
library(ncdf4)
library(raster)
library(proj4)
library(rgdal) # needed to read in boulder shapefile
library(daymetr)
library(ggplot2)
library(broom)
library(tidyverse)
library(ggspatial)
library(rasterVis) # for levelplot
```

```{r load-boulder-city-outline}
boulder_outline <- readOGR("data/BoulderCityLimits/BoulderCityLimits.shp")
plot(boulder_outline)
extent(boulder_outline)
crs(boulder_outline)
```

```{r convert-crs-city-outline}
# # create a region of interest, example is: c(top left lat, top left long, bottom right lat, bottom right long)
# roi <- c(40.097, -105.304, 40.002, -105.175)
# # Define the variables for data extraction
# start_year <- 2012
# end_year <- 2012
# climate_var <- "tmin" # c("tmin", "tmax") or "ALL"
# data_path <- "data/daymet/boulder-roi-40.097-105.304"
data_path2 <- "/Volumes/UrbanClimatePheno/daymet" # i think the file is too large for my computer, switch to an external harddrive for raw data
# 
# # use daymet API to download the data in the region of interest
# download_daymet_ncss(location = roi,
#                       start = start_year,
#                       end = end_year,
#                       frequency = "monthly",
#                       param = climate_var,
#                       path = data_path2, 
#                       silent = F)
tmin.mon.s <- raster::stack(paste0(data_path2,"/tmin_monavg_2012_ncss.nc"))

# > crs(tmin.mon.s)
# CRS arguments:
#  +proj=lcc +lon_0=-100 +lat_0=42.5 +x_0=0 +y_0=0 +lat_1=25 +a=60 +rf=6378137 +lat_2=45 

# try reprojecting to a meters crs
tmin.mon.s_repr <- projectRaster(tmin.mon.s, crs="+proj=lcc +lat_1=25.00 +lat_2=60.00 +lat_0=42.5 +lon_0=-100.00 +x_0=0 +y_0=0 +ellps=GRS80 +units=m +no_defs")

# Reproject the city polygons onto the raster brick's coord ref system
boulder_outline_reproj3 <- spTransform(boulder_outline,
                                crs(proj4string(tmin.mon.s_repr)))

```
It's convoluted, but the overlay seems to work by creating the city outline using this round a bout way to get the extent and crs below.

crs(boulder_outline_reproj3)

CRS arguments:
 +proj=lcc +lat_1=25.00 +lat_2=60.00 +lat_0=42.5 +lon_0=-100.00 +x_0=0 +y_0=0 +ellps=GRS80 +units=m
+no_defs 

extent(boulder_outline_reproj3)
> extent(boulder_outline_reproj3)
class       : Extent 
xmin        : -431996.4 
xmax        : -421564.2 
ymin        : -255924.2 
ymax        : -241673.5 


```{r add-southern-tile-2013}

# create a region of interest, example is: c(top left lat, top left long, bottom right lat, bottom right long)
roi <- c(40.09, -105.32, 39.95, -105.17)
# Define the variables for data extraction
start_year <- 2013
end_year <- 2013
climate_var <- "tmin" # c("tmin", "tmax") or "ALL"
data_path2 <- "/Volumes/UrbanClimatePheno/daymet" # i think the file is too large for my computer, switch to an external harddrive for raw data

# # use daymet API to download the data in the region of interest
#  download_daymet_tiles(location = roi,
#                       start = start_year,
#                       end = end_year,
#                       tile = NULL,
#                       param = climate_var,
#                       path = data_path2,
#                       silent = F)
 
tmin13.s.n <- raster::stack(paste0(data_path2,"/tmin_2013_11738.nc"))
tmin13.s.s <- raster::stack(paste0(data_path2,"/tmin_2013_11558.nc"))

plot(tmin13.s.n)
plot(tmin13.s.s)
print(tmin13.s.n)
print(tmin13.s.s)

# try plotting on tile
plot(tmin13.s.n[[1]])
plot(boulder_outline_reproj3, add = TRUE)
plot(tmin13.s.s[[1]])
plot(boulder_outline_reproj3, add = TRUE)


# set extents
l.extent.n <- c(-439000, -419000, -279000, -240000)

# trim extent of raster around city buffer
tmin13.s.n.c <- crop(tmin13.s.n, l.extent.n)
extent(tmin13.s.n.c)

# southern tile cropped to boulder to get idea of extent

test_crop <- crop(tmin13.s.s, boulder_outline_reproj3)
extent(test_crop)
plot(test_crop[[2]])
plot(boulder_outline_reproj3, add  = T)

l.extent.s <- c(-439000, -419000,-275500,-257500)
l.extent.s2 <- c( -439000, -419000,-260000, -241673)
# trim extent of raster around city buffer
tmin13.s.s.c <- crop(tmin13.s.s, l.extent.s2)
extent(tmin13.s.s.c)

# check plot
plot(tmin13.s.s.c[[2]])
plot(boulder_outline_reproj3, add  = T)

# try to overlap rasters with a single layer
test_merge <- merge(tmin13.s.n.c[[2]], tmin13.s.s.c[[2]])
plot(test_merge)
plot(boulder_outline_reproj3, add = T)

# try to merge raster stack
stack_merge <- merge(tmin13.s.n.c, tmin13.s.s.c)
plot(stack_merge)
plot(stack_merge[[2]])
plot(boulder_outline_reproj3, add = T)




```


```{r jan-12-days}
# trim to first 12 days of the month
tmin13.jan <- stack_merge[[1:12]]

# convert from C to F
r5 = (tmin13.jan*9)/5 + 32

summary(r5) # range is from -5.8 to 31

levelplot(r5, 
          margin=FALSE,                       
          colorkey=list(
            space='bottom',                   
            labels=list(at=-6:32, font=4),
            axis.line=list(col='black'),
            width=0.75
          ),
          main=list(label="Daymet Derived Min Temps (F) for Boulder, CO (Jan 1-12, 2013)", cex=.8),
          scales=list(draw=FALSE),            
          col.regions=colorRampPalette(rev(c('dark red','white','dark blue'))),                   
          at=seq(-6, 32, len=101),
          names.attr= c("Jan 1, 2013", "Jan 2, 2013", "Jan 3, 2013", "Jan 4, 2013", "Jan 5, 2013", "Jan 6, 2013", "Jan 7, 2013", "Jan 8, 2013", "Jan 9, 2013", "Jan 10, 2013", "Jan 11, 2013", "Jan 12, 2013")) +           
  layer(sp.polygons(boulder_outline_reproj3, lwd=1))


```

```{r apr-12-days}
# trim to first 12 days of the month
tmin13.apr <- stack_merge[[91:102]]

# convert from C to F
r6 = (tmin13.apr*9)/5 + 32

summary(r6) # range is from -3.1 to 49

levelplot(r6, 
          margin=FALSE,                       
          colorkey=list(
            space='bottom',                   
            labels=list(at=-4:49, font=4),
            axis.line=list(col='black'),
            width=0.75
          ),
          main=list(label="Daymet Derived Min Temps (F) for Boulder, CO (Apr 1-12, 2013)", cex=.8),
          scales=list(draw=FALSE),            
          col.regions=colorRampPalette(rev(c('dark red','white','dark blue'))),                   
          at=seq(-4, 49, len=101),
          names.attr= c("Apr 1, 2013", "Apr 2, 2013", "Apr 3, 2013", "Apr 4, 2013", "Apr 5, 2013", "Apr 6, 2013", "Apr 7, 2013", "Apr 8, 2013", "Apr 9, 2013", "Apr 10, 2013", "Apr 11, 2013", "Apr 12, 2013")) +           
  layer(sp.polygons(boulder_outline_reproj3, lwd=1))


```

```{r jul-12-days}
# trim to first 12 days of the month
tmin13.jul <- stack_merge[[182:193]]

# convert from C to F
r7 = (tmin13.jul*9)/5 + 32

summary(r7) # range is from 41 to 66

levelplot(r7, 
          margin=FALSE,                       
          colorkey=list(
            space='bottom',                   
            labels=list(at=41:66, font=4),
            axis.line=list(col='black'),
            width=0.75
          ),
          main=list(label="Daymet Derived Min Temps (F) for Boulder, CO (Jul 1-12, 2013)", cex=.8),
          scales=list(draw=FALSE),            
          col.regions=colorRampPalette(rev(c('dark red','white','dark blue'))),                   
          at=seq(41, 66, len=101),
          names.attr= c("Jul 1, 2013", "Jul 2, 2013", "Jul 3, 2013", "Jul 4, 2013", "Jul 5, 2013", "Jul 6, 2013", "Jul 7, 2013", "Jul 8, 2013", "Jul 9, 2013", "Jul 10, 2013", "Jul 11, 2013", "Jul 12, 2013")) +           
  layer(sp.polygons(boulder_outline_reproj3, lwd=1))

```

```{r oct-12-days}
# trim to first 12 days of the month
tmin13.oct <- stack_merge[[274:285]]

# convert from C to F
r8 = (tmin13.oct*9)/5 + 32

summary(r8) # range is from 23.9 to 50.9

levelplot(r8, 
          margin=FALSE,                       
          colorkey=list(
            space='bottom',                   
            labels=list(at=23:51, font=4),
            axis.line=list(col='black'),
            width=0.75
          ),
          main=list(label="Daymet Derived Min Temps (F) for Boulder, CO (Oct 1-12, 2013)", cex=.8),
          scales=list(draw=FALSE),            
          col.regions=colorRampPalette(rev(c('dark red','white','dark blue'))),                   
          at=seq(23, 51, len=101),
          names.attr= c("Oct 1, 2013", "Oct 2, 2013", "Oct 3, 2013", "Oct 4, 2013", "Oct 5, 2013", "Oct 6, 2013", "Oct 7, 2013", "Oct 8, 2013", "Oct 9, 2013", "Oct 10, 2013", "Oct 11, 2013", "Oct 12, 2013")) +           
  layer(sp.polygons(boulder_outline_reproj3, lwd=1))

```


```{r ggplot-works}
# https://stackoverflow.com/questions/33227182/how-to-set-use-ggplot2-to-map-a-raster

# install.packages("ggthemes")
library(ggthemes) # theme_map()
# install.packages("viridis")
library(viridis)  # better colors for everyone
library(scales)

r <- stack_merge[[2]]

test_spdf <- as(r, "SpatialPixelsDataFrame")
test_df <- as.data.frame(test_spdf)
colnames(test_df) <- c("value", "x", "y")


ggplot() +  
  geom_tile(data=test_df, aes(x=x, y=y, fill=value), alpha=0.8) + 
  geom_polygon(data=boulder_outline_reproj3, aes(x=long, y=lat, group=group), 
               fill=NA, color="grey50", size=0.25) +
  scale_fill_viridis() +
  coord_equal() +
  theme_map() +
  theme(legend.position="bottom") +
  theme(legend.key.width=unit(2, "cm"))




```

Another source for plotting with ggplot
https://datacarpentry.org/r-raster-vector-geospatial/aio.html

ggplot() + 
  geom_sf(data = aoi_boundary_HARV, size = 3, color = "black", fill = "cyan1") + 
  ggtitle("AOI Boundary Plot") + 
  coord_sf()


```{r test-2018}
# create a region of interest, example is: c(top left lat, top left long, bottom right lat, bottom right long)
roi <- c(40.097, -105.304, 40.002, -105.175)
# Define the variables for data extraction
start_year <- 2018
end_year <- 2018
climate_var <- "tmin" # c("tmin", "tmax") or "ALL"
data_path <- "data/daymet/boulder-roi-40.097-105.304"
data_path2 <- "/Volumes/UrbanClimatePheno/daymet" # i think the file is too large for my computer, switch to an external harddrive for raw data

# use daymet API to download the data in the region of interest
download_daymet_ncss(location = roi,
                      start = start_year,
                      end = end_year,
                      frequency = "daily",
                      param = climate_var,
                      path = data_path2,
                      silent = F)
# open raster
tmin.2018 <- raster::stack(paste0(data_path2,"/test/tmin_monavg_2018_ncss.nc"))

# try reprojecting to a meters crs
tmin.mon.s_repr <- projectRaster(tmin.mon.s, crs="+proj=lcc +lat_1=25.00 +lat_2=60.00 +lat_0=42.5 +lon_0=-100.00 +x_0=0 +y_0=0 +ellps=GRS80 +units=m +no_defs")

# Define the variables for data extraction
start_year <- 2018
end_year <- 2018
climate_var <- "tmin" # c("tmin", "tmax") or "ALL"
data_path2 <- "/Volumes/UrbanClimatePheno/daymet" # i think the file is too large for my computer, switch to an external harddrive for raw data

# use daymet API to download the data in the region of interest
 download_daymet_tiles(location = roi,
                      start = start_year,
                      end = end_year,
                      tile = NULL,
                      param = climate_var,
                      path = data_path2,
                      silent = F)
 
tmin18.s.n <- raster::stack(paste0(data_path2,"/tmin_2018_11738.nc"))
plot(tmin18.s.n)

```

