---
title: "04_bmc_data_exploration_yearly"
author: "Deidre Jaeger"
date: "1/10/2019"
output: html_document
---
TO dos

- apply to full span of years!

- get a DEM for boulder area to compare (or underlay on the temp map- see NEON tutorial)

USC00050848	BOULDER	39.9919	-105.267

Questions:
Is there a spatial pattern to temp variability? 
Is there a detectable urban heat island effect? 
How does this compare to the overall climate change signal? (using landsat?)

Analysis plan
1) define an area in the city and outside (or a transect), compare the mean yearly temps for a section inside the city to a section outside the city see if significantly different for a 30 year comparison- what is the difference?

Create a study transcect (or 3?) and crop values, convert to a tidy dateframe so can look at profiles and compare quantitatively

Carol suggests using NLCD or tree cover data to show the difference in vegetation... or could also look at NDVI? 

```{r install-packages}
library(ncdf4)
library(raster)
library(proj4)
library(rgdal) # needed to read in boulder shapefile
library(daymetr)
library(ggplot2)
library(broom)
library(tidyverse)
library(ggspatial)
library(rasterVis) # for levelplot
```

```{r load-boulder-city-outline}
# load boulder shapefile
boulder_outline <- readOGR("data/BoulderCityLimits/BoulderCityLimits.shp")
# set path to data on external harddrive
data_path2 <- "/Volumes/UrbanClimatePheno/daymet"
# load a raster with monthly aggregations
tmin.mon.s <- raster::stack(paste0(data_path2,"/tmin_monavg_2012_ncss.nc"))
# reproject coordinate reference system
tmin.mon.s_repr <- projectRaster(tmin.mon.s, crs="+proj=lcc +lat_1=25.00 +lat_2=60.00 +lat_0=42.5 +lon_0=-100.00 +x_0=0 +y_0=0 +ellps=GRS80 +units=m +no_defs")
# Reproject the city polygons onto the raster brick's coord ref system
boulder_outline_reproj3 <- spTransform(boulder_outline,
                                crs(proj4string(tmin.mon.s_repr)))
```



```{r download-2014-daily-data}

# create a region of interest, example is: c(top left lat, top left long, bottom right lat, bottom right long)
roi <- c(40.09, -105.32, 39.95, -105.17)
# Define the variables for data extraction
start_year <- 2014
end_year <- 2014
climate_var <- "tmin" # c("tmin", "tmax") or "ALL"
data_path2 <- "/Volumes/UrbanClimatePheno/daymet" # i think the file is too large for my computer, switch to an external harddrive for raw data

# use daymet API to download the data in the region of interest
 download_daymet_tiles(location = roi,
                      start = start_year,
                      end = end_year,
                      tile = NULL,
                      param = climate_var,
                      path = data_path2,
                      silent = F)
 
tmin14.s.n <- raster::stack(paste0(data_path2,"/tmin_2014_11738.nc"))
tmin14.s.s <- raster::stack(paste0(data_path2,"/tmin_2014_11558.nc"))

plot(tmin14.s.n)
plot(tmin14.s.s)
print(tmin14.s.n)
print(tmin14.s.s)

# try plotting one tile
plot(tmin14.s.n[[1]])
plot(boulder_outline_reproj3, add = TRUE)
plot(tmin14.s.s[[1]])
plot(boulder_outline_reproj3, add = TRUE)


# set extents, could also use extent(xmin, xmax, ymin, ymax)
l.extent.n <- c(-439000, -419000, -279000, -240000)
# l.extent.s <- c(-439000, -419000,-275500,-257500)
l.extent.s2 <- c( -439000, -419000,-260000, -241673)

# trim northern extent of raster around city buffer
tmin14.s.n.c <- crop(tmin14.s.n, l.extent.n)
extent(tmin14.s.n.c)

# trim southern extent of raster around city buffer
tmin14.s.s.c <- crop(tmin14.s.s, l.extent.s2)
extent(tmin14.s.s.c)


# merge raster stack for 2014 daily min temp values
stack_merge <- merge(tmin14.s.n.c, tmin14.s.s.c)

```

```{r monthly-data-2014}
# https://www.rdocumentation.org/packages/raster/versions/2.8-4/topics/calc

# use overlay() for single raster math, use calc() for raster stacks 
# If x is a RasterStack or RasterBrick, fun should operate on a vector of values (one vector for each cell). calc returns a RasterLayer if fun returns a single value (e.g. sum) and it returns a RasterBrick if fun returns more than one number, e.g., fun=quantile

# calculate the mean min temp for each pixel in a given month
jan_mean <- mean(stack_merge[[1:31]])
feb_mean <- mean(stack_merge[[32:59]])
mar_mean <- mean(stack_merge[[60:90]])
apr_mean <- mean(stack_merge[[91:120]])
may_mean <- mean(stack_merge[[121:151]])
jun_mean <- mean(stack_merge[[152:181]])
jul_mean <- mean(stack_merge[[182:212]])
aug_mean <- mean(stack_merge[[213:243]])
sep_mean <- mean(stack_merge[[244:273]])
oct_mean <- mean(stack_merge[[274:304]])
nov_mean <- mean(stack_merge[[305:334]])
dec_mean <- mean(stack_merge[[335:365]])

# stack the monthly rasters
month2014_ave <- raster::stack(jan_mean, feb_mean, mar_mean, apr_mean, may_mean, jun_mean, jul_mean, aug_mean, sep_mean, oct_mean, nov_mean, dec_mean)

plot(month2014_ave)

# convert to from C to F
month2014_ave = (month2014_ave*9)/5 + 32


```

```{r yearly-2014}

# calc average min temp for 2014
tmin_2014 <- calc(month2014_ave, mean)

# summary
summary(tmin_2014)


levelplot(tmin_2014, 
          margin=FALSE,                       
          colorkey=list(
            space='bottom',                   
            labels=list(at=30:39, font=4),
            axis.line=list(col='black'),
            width=0.75
          ),
          main=list(label="Daymet Derived mean Min Temps (F) for Boulder, CO (2014)", cex=.8),
          scales=list(draw=FALSE),            
          col.regions=colorRampPalette(rev(c('dark red','white','dark blue'))),                   
          at=seq(30, 39, len=20)) +
          # names.attr= c("Jan, 2014", "Feb, 2014", "Mar, 2014", "Apr, 2014", "May, 2014", "Jun, 2014", "Jul, 2014", "Aug, 2014", "Sep, 2014", "Oct, 2014", "Nov, 2014", "Dec, 2014")) +           
  layer(sp.polygons(boulder_outline_reproj3, lwd=2))


```


